### 2018-06-13 CS761-RandomizedAlgorithms/assignments/a2.tex
partial q2

\begin{document}
%Problem 1
\section{}
\paragraph{}
Let $v_1, \dots, v_m \in \{0,1\}^\ell$ be a set of vectors such that any subset of size at most $k$ is linearly independent modulo $2$. We denote the $j$-th entry of $v_i$ by $v_{i,j}$. Let $u$ be chosen uniformly at random from $\{0,1\}^\ell$. Let $X_i$ be the random variable defined by
$$X_i = (\sum_{j=1}^\ell v_{i,j}u_j) \mod 2.$$
We want to show that $X_1, \dots, X_m$ are $k$-wise independent random bits. That is to say, for all $I \subseteq [m]$ with $1\leq |I| \leq k$, and $x \in \{0,1\}^m$,
$$Pr(\bigcap_{i\in I} X_i = x_i) = \Pi_{i\in I} Pr(X_i = x_i).$$
We begin by observing that if the vector $v_i$ is the $0$ vector for any $i \in I$ then the set of vectors $V := \{v_i : i \in I\}$ fails to be linearly independent (modulo $2$), which contradicts our hypothesis. Hence for all $i \in I$, there exists $j \in [\ell]$ such that $v_{i,j} \neq 0$. Further from linear independence we observe that $\ell \geq k$.
\paragraph{}
On the one hand, for any $i \in I$, $v_i \neq 0$ and so
$$Pr(X_i = x_i) = \frac{1}{2}$$
since $X_i = x_i$ if and only if
$$P_i := \sum_{j \in s(v_i)} u_j = \begin{cases} \text{odd}, &\text{if $x_i = 1$} \\
\text{even}, &\text{otherwise}
\end{cases}$$
where $s(v_i) := \{j \in [\ell] : v_{i,j} \neq 0\}$. Note that since $s(v_i) \neq \emptyset$ and $u$ is chosen uniformly at random, the parity $P_i$ is chosen uniformly at random. Therefore
$$\Pi_{i \in I} Pr(X_i = x_i) = \Pi_{i \in I} \frac{1}{2} = \frac{1}{2^{|I|}}.$$
\paragraph{}
Now on the other hand, $\bigcap_{i \in I} X_i = x_i$ happens if and only if 
$$ Vu = x_I$$
where we define the matrix $V$ with $|I|$ rows and $\ell$ columns so that $V_{i,j} = v_{i,j}$, and $x_I$ is the vector $x$ restricted to entries indexed by $I$. Here we are taking all operations in the field $\F_2$. Since the rank of $V$ is $|I|$ we can find by Gaussian elimination a permuation matrix $E$ (after possibly permuting the columns and associated indices of $u$, without loss of generality) so that
$$EV = \begin{bmatrix}
I & B
\end{bmatrix}$$
for some binary matrix $B$ with $\ell - |I|$ columns. Then
$$\begin{bmatrix} I & B \end{bmatrix}u = Ex_I$$
and hence the choices of entries for $u_1, \dots, u_{|I|}$ are uniquely fixed given the choices of entries for $u_{|I| + 1}, \dots, u_{\ell}$. So there are $2^{\ell - |I|}$ choices for $u$ satisfying our matrix equation. There are $2^\ell$ total choices for $u$ overall, and so we have
$$Pr(\bigcap_{i \in I} X_i = x_i) = \frac{2^{\ell-|I|}}{2^\ell}  = \frac{1}{2^{|I|}} = \Pi_{i \in I} Pr(X_i = x_i)$$
as desired. $\blacksquare$
%Problem 2
\newpage
\section{}

%Problem 3
\newpage
\section{}

>>>>

\begin{document}
%Problem 1
\section{}
\paragraph{}
Let $v_1, \dots, v_m \in \{0,1\}^\ell$ be a set of vectors such that any subset of size at most $k$ is linearly independent modulo $2$. We denote the $j$-th entry of $v_i$ by $v_{i,j}$. Let $u$ be chosen uniformly at random from $\{0,1\}^\ell$. Let $X_i$ be the random variable defined by
$$X_i = (\sum_{j=1}^\ell v_{i,j}u_j) \mod 2.$$
We want to show that $X_1, \dots, X_m$ are $k$-wise independent random bits. That is to say, for all $I \subseteq [m]$ with $1\leq |I| \leq k$, and $x \in \{0,1\}^m$,
$$Pr(\bigcap_{i\in I} X_i = x_i) = \Pi_{i\in I} Pr(X_i = x_i).$$
We begin by observing that if the vector $v_i$ is the $0$ vector for any $i \in I$ then the set of vectors $V := \{v_i : i \in I\}$ fails to be linearly independent (modulo $2$), which contradicts our hypothesis. Hence for all $i \in I$, there exists $j \in [\ell]$ such that $v_{i,j} \neq 0$. Further from linear independence we observe that $\ell \geq k$.
\paragraph{}
On the one hand, for any $i \in I$, $v_i \neq 0$ and so
$$Pr(X_i = x_i) = \frac{1}{2}$$
since $X_i = x_i$ if and only if
$$P_i := \sum_{j \in s(v_i)} u_j = \begin{cases} \text{odd}, &\text{if $x_i = 1$} \\
\text{even}, &\text{otherwise}
\end{cases}$$
where $s(v_i) := \{j \in [\ell] : v_{i,j} \neq 0\}$. Note that since $s(v_i) \neq \emptyset$ and $u$ is chosen uniformly at random, the parity $P_i$ is chosen uniformly at random. Therefore
$$\Pi_{i \in I} Pr(X_i = x_i) = \Pi_{i \in I} \frac{1}{2} = \frac{1}{2^{|I|}}.$$
\paragraph{}
Now on the other hand, $\bigcap_{i \in I} X_i = x_i$ happens if and only if 
$$ Vu = x_I$$
where we define the matrix $V$ with $|I|$ rows and $\ell$ columns so that $V_{i,j} = v_{i,j}$, and $x_I$ is the vector $x$ restricted to entries indexed by $I$. Here we are taking all operations in the field $\F_2$. Since the rank of $V$ is $|I|$ we can find by Gaussian elimination a permuation matrix $E$ (after possibly permuting the columns and associated indices of $u$, without loss of generality) so that
$$EV = \begin{bmatrix}
I & B
\end{bmatrix}$$
for some binary matrix $B$ with $\ell - |I|$ columns. Then
$$\begin{bmatrix} I & B \end{bmatrix}u = Ex_I$$
and hence the choices of entries for $u_1, \dots, u_{|I|}$ are uniquely fixed given the choices of entries for $u_{|I| + 1}, \dots, u_{\ell}$. So there are $2^{\ell - |I|}$ choices for $u$ satisfying our matrix equation. There are $2^\ell$ total choices for $u$ overall, and so we have
$$Pr(\bigcap_{i \in I} X_i = x_i) = \frac{2^{\ell-|I|}}{2^\ell}  = \frac{1}{2^{|I|}} = \Pi_{i \in I} Pr(X_i = x_i)$$
as desired. $\blacksquare$
%Problem 2
\newpage
\section{}
\paragraph{}
Suppose we have $n$ items to be hashed into $n$ bins. Let $h$ be a hash function chosen from a $k$-universal family. We $\textit{overload}$ the notation ${n \choose k}$ to mean both the number of ways to choose $k$ elements from $n$, and the set of subsets of $[n]$ of size $k$. Which version we use will always be clear from context. Define a $k$-collision to be the event that $k$ distinct items are hashed to the same bin. For a $\textit{set}$ $S \in {n\choose k}$ let $X_S$ be the random indicator variable which is $1$ when $h(x) = h(y)$ for all $x,y \in S$, and $X_S$ is $0$ otherwise.
\paragraph{}
Let $X$ be a random variable counting the number of $k$-collisions when our $n$ items are hased using $h$ chosen from a $k$-universal family. Then $X = \sum_{S \in {n\choose k}} X_S$ and so we have
$$E[X] = E[\sum_{S \in {n\choose k}} X_S] = \sum_{S \in {n\choose k}} E[X_S] = \sum_{S \in {n \choose k}} Pr(|h(S)| = 1) \leq  \sum_{S \in {n \choose k}} \frac{1}{n^{k-1}}$$
where the inequality follows since $h$ is a $k$-universal hash function, and $S$ is a set of size $k$.
%Problem 3
\newpage
\section{}
